---
title: "Untitled"
author: "Gilad Ravid"
date: "11/18/2021"
output:
  html_document: default
  pdf_document: default
---
# One Dataset, Visualized 25 Ways
## This is what happens when you let the data ramble on and on and on…
based on https://flowingdata.com/2017/01/24/one-dataset-visualized-25-ways/

# “Let the data speak.”

```{r setup}
library(dplyr)
library(ggplot2)
library(tidyr)
#remotes::install_github("yaweige/ggpcp", build_vignettes = TRUE)
library(ggpcp)
```
```{r dataPrep}
# https://data.worldbank.org/indicator/SP.DYN.LE00.FE.IN

setwd("~/Dropbox (BGU)/courses/InfoVis_2022a")

le <- read.csv("datasets/LifeExpec.csv",header = TRUE)
leM <- read.csv("datasets/LifeExpecMale.csv",header = TRUE)
leF <- read.csv("datasets/LifeExpecFemale.csv",header = TRUE)

names(le)[3:62]<-substr(names(le)[3:62],2,6)
names(leM)[3:62]<-substr(names(leM)[3:62],2,6)
names(leF)[3:62]<-substr(names(leF)[3:62],2,6)

leLong<- le %>% pivot_longer(cols=3:62,names_to="year",values_to = "total")
leMLong<- leM %>% pivot_longer(cols=3:62,names_to="year",values_to = "male")
leFLong<- leF %>% pivot_longer(cols=3:62,names_to="year",values_to = "female")

leLong$male<-leMLong$male
leLong$female<-leFLong$female

leLong$year<-as.numeric(leLong$year)

leDiff <- leLong %>% 
  group_by(Country.Code) %>% 
  arrange(year) %>%
  mutate(diff=(total - lag(total,n=1,default=first(total)))*100/lag(total,n=1,default=first(total)),
         prevTot=lag(total,n=1,default=first(total)))

leDiff_2019<-leDiff[leDiff$year==2019,]
meanTotal2019<-mean(leDiff_2019$total,na.rm=T)
leDiff_2019<-leDiff_2019 %>% 
  mutate(AvgDiff=(total - meanTotal2019),pos=AvgDiff>0)
```

\newpage

```{r p1}
p1<-ggplot(leLong,aes(x=year,y=total,group=Country.Code))+geom_line()+ 
  ggtitle("Life Expectancy at Birth, 1960-2019 \n Each line represents a country") +
  xlab("") + ylab("")
p1
```

## Looks like spaghetti

\newpage

```{r p2,fig.width = 25, fig.height= 30}
p2<-ggplot(leLong,aes(x=year,y=total))+
  geom_line()+facet_wrap(~Country.Name,ncol=20)+
  ggtitle("Life Expectancy at Birth, 1960-2019") + 
  theme(strip.text.x = element_text(size = 6),axis.title = element_blank(), 
        axis.text.y = element_blank(),axis.ticks = element_blank(),
        axis.text.x = element_blank(),axis.ticks.x = element_blank(),
        panel.border = element_blank(),panel.background=element_blank(),
        panel.grid.major = element_blank(),panel.grid.minor = element_blank())
p2
```

## Shows countries separately so that lines don’t overlap

\newpage

```{r p3,fig.width = 25, fig.height= 30}
p3<-ggplot(leDiff_2019,aes(x=reorder(Country.Name, total),y=total))+
  geom_bar(stat = "identity")+
  coord_flip()+
  ggtitle("Life Expectancy at Birth,2019")+xlab("")+ylab("")
p3
```

## No surprises

\newpage

```{r p4,fig.width = 25, fig.height= 30}
p4<-ggplot(leDiff_2019)+
  geom_segment(aes(x=male,xend=female,y=reorder(Country.Name, total),
                   yend=reorder(Country.Name, total)))+
  ggtitle("Life Expectancy at Birth For Males and Females,2019")+
  xlab("")+
  ylab("")
p4
```

## Focus on the the difference between the two genders, with comparison across countries

\newpage

```{r p5,fig.width = 25, fig.height= 30}
p5<-ggplot(leLong,aes(x=year))+
  geom_line(aes(y=male))+
  geom_line(aes(y=female))+
  facet_wrap(~Country.Name,ncol=10)+
  ggtitle("Life Expectancy at Birth for Male and Female, 1960-2019") + 
  theme(strip.text.x = element_text(size = 4),axis.title = element_blank(), 
        axis.text.y = element_blank(), axis.text.x = element_blank(),
        panel.border = element_blank())
p5
```

## Focus on difference between male and female over time

\newpage

```{r p6,fig.width = 25, fig.height= 30}
names(le)[4]<-"X1961"
names(le)[62]<-"X2019"
p6<-ggplot(le)+
  geom_segment(aes(x=X1961,xend=X2019,y=reorder(Country.Name, X2019),
                   yend=reorder(Country.Name, X2019)))+
  ggtitle("Life Expectancy at Birth 1961 vs. 2019")+
  xlab("")+
  ylab("")
p6
```

## A focus the change between two time periods instead of the fluctuations

\newpage

```{r p7}
p7<-ggplot(leDiff_2019,aes(x=male,y=female))+
  geom_point()+
  ggtitle("Life Expectancy at Birth for Males and Female, 2019")+
  xlab("Male")+
  ylab("Female")
p7
```

## Comparison between the two, in a more compact space
```{r p7b}
p7b<-ggplot(leDiff_2019) + 
  geom_pcp(aes(vars=vars(male,female,total)))
p7b
```
\newpage

```{r p8}
p8<-ggplot(leLong,aes(x=male,y=female,group=Country.Code))+
  geom_line()+
  theme(legend.position = "none")+
  ggtitle("Life Expectancy at Birth for Males vs. Female, 1960-2019")+
  xlab("Male")+
  ylab("Female")
p8
```

## Shows changes over time, although not super clear with this dataset
```{r p8a}
p8<-ggplot(leLong[1:500,],aes(x=male,y=female,group=Country.Code))+
  geom_line(aes(colour=Country.Code))+
  theme(legend.position = "none")+
  ggtitle("Life Expectancy at Birth for Males vs. Female, 1960-2019")+
  xlab("Male")+
  ylab("Female")
p8
```

\newpage



\newpage

```{r p9,fig.width = 25, fig.height= 30}
p9<-ggplot(leLong,aes(x=male,y=female))+
  geom_line()+
  facet_wrap(~Country.Name,ncol=10)+
  ggtitle("Life Expectancy at Birth for Male vs. Female, 1960-2019") + 
  theme(strip.text.x = element_text(size = 4),axis.title = element_blank(), 
        axis.text.y = element_blank(), axis.text.x = element_blank(),
        panel.border = element_blank())
p9

```

## Unjumbles the lines for better visibility, but harder to see overall patterns because of size

\newpage

```{r p10}
p10<-ggplot(leDiff,aes(x=year,y=diff,group=Country.Code))+
  geom_line()+
  ggtitle("Life Expectancy at Birth Year over Year Percentage Change, 1960-2019")+
  xlab("")+
  ylab("")
p10
```

## Places emphasis on the annual changes instead of the actual values

\newpage

```{r p11,fig.width = 25, fig.height= 30}
p11<-ggplot(leDiff,aes(x=year,y=diff))+
  geom_line()+
  facet_wrap(~Country.Name,ncol=10)+
  ggtitle("Life Expectancy at Birth % Change from Previous Year, 1960-2019") + 
  theme(strip.text.x = element_text(size = 4),axis.title = element_blank(), 
        axis.text.y = element_blank(), axis.text.x = element_blank(),
        panel.border = element_blank())
p11
```

## Most patterns obscured by large values

\newpage

```{r p12}
p12<-ggplot(leDiff[leDiff$Country.Code %in% c("ISR","KHM","RWA","TLS","IRN","ZMB")
                   & leDiff$year>1969 & leDiff$year<1991,],aes(x=year,y=1))+
  geom_point(aes(colour=diff,size=5))+
  facet_wrap(~Country.Name,ncol=3)+
  scale_colour_gradient2(low = "#870ec2",mid="#ffffff",high = "#3c930f",space = "Lab",
                         na.value = "grey50",guide = "colourbar",aesthetics = "colour")+
  theme(panel.border = element_blank(), panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(), axis.line = element_line(colour = "black"),
        axis.title = element_blank(), axis.text.y = element_blank())+
  ggtitle("Life Expectancy at Birth Change from Previous Year, 1960-2019")
p12
```

# Focusing on a handful of countries with larger fluctuations

\newpage

```{r p13,fig.width = 25, fig.height= 30}
p13<-ggplot(leDiff_2019,aes(x=1,y=1,size=total))+ 
  geom_point()+
  facet_wrap(~Country.Name,ncol=10)+
  theme(strip.text.x = element_text(size = 4),axis.title = element_blank(), 
        axis.text.y = element_blank(),axis.ticks = element_blank(), 
        axis.text.x = element_blank(),axis.ticks.x = element_blank(),
        panel.border = element_blank(), panel.grid.major = element_blank(),
        panel.grid.minor = element_blank())+
  ggtitle("Life Expectancy at Birth in 2019")
p13
```

## One bubble for each country in 2019

\newpage

```{r p14 ,fig.width = 25, fig.height= 30}
p14<-ggplot(leLong,aes(x=year,y=Country.Name,fill=total))+
  geom_tile()+
  ggtitle("Life Expectancy at Birth in 1960-2019")
p14
```

## Shows everything in a more compact space, but with less distinct separation for countries

\newpage

```{r p15}
p15<-ggplot(leLong,aes(x=year,y=total,group=Country.Code))+
  geom_line(color="orange",alpha=0.2)+
  coord_polar()+
  ggtitle("Life Expectancy at Birth in 1960-2019")+
  xlab("")+
  ylab("")
p15
```

## Mostly for show, but some sense of distribution

\newpage

```{r p16}
p16<-ggplot(leLong,aes(x=as.factor(year),y=total))+
  geom_boxplot()+
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))+
  ggtitle("Life Expectancy at Birth\nDistribution of Coutries, 1960-2019")+
  xlab("")+
  ylab("Years")
p16
```

## The classics, focusing on spread

\newpage

```{r p17,fig.width = 25, fig.height= 30}
p17<-ggplot(leLong,aes(x=total))+
  geom_histogram()+
  facet_wrap(~year,ncol=5)+
  ggtitle("Life Expectancy at Birth\nDistribution of Coutries, 1960-2019")+
  xlab("years")+
  ylab("Count")
p17
```

## Focus on distributions for each year instead of individual countries

\newpage

```{r p18,fig.width = 25, fig.height= 30}
p18<-ggplot(leLong,aes(x=total))+
  geom_density()+
  facet_wrap(~year,ncol=5)+
  ggtitle("Life Expectancy at Birth\nDistribution of Coutries, 1960-2019")+
  xlab("years")+
  ylab("Density")
p18
```

## Smooth versions of histograms, but a bit more difficult to decipher

\newpage

```{r p19}
p19<-ggplot(leLong[leLong$year>2009,],aes(x=as.factor(year),y=total))+
  geom_dotplot(binaxis='y', stackdir='center',stackratio=0.9, dotsize=0.2)+
  theme(axis.text.x = element_text(angle = 45, vjust = 0.5, hjust=1))+
  ggtitle("Life Expectancy at Birth\nDistribution of Coutries, 1960-2019")+
  xlab("")+
  ylab("")
p19
```

## Get an idea of spread and still get some individuality

\newpage

```{r p20}
p20<-ggplot(leLong[leLong$year>2009,],aes(x=as.factor(year),y=total))+
  geom_violin(trim=FALSE)+
  geom_dotplot(binaxis='y', stackdir='center',stackratio=0.9, dotsize=0.2)+
  theme(axis.text.x = element_text(angle = 45, vjust = 0.5, hjust=1))+
  ggtitle("Life Expectancy at Birth\nDistribution of Coutries, 1960-2019")+
  xlab("")+
  ylab("")
p20
```

## Combining charts is an option too

\newpage

```{r p21,fig.width = 25, fig.height= 30}
p21<-ggplot(leLong,aes(x=total))+
  geom_histogram()+geom_histogram(aes(y=-..count..))+
  coord_flip()+facet_wrap(~year,ncol=5)+
  scale_y_continuous(labels=c("50","25","0","25","50"), breaks=c(-25,-12.5,0,12.5,25), 
                     limits=c(-50,50))+
  ggtitle("Life Expectancy at Birth\nDistribution of Coutries, 1960-2019")+
  xlab("Years")+
  ylab("")
p21
```

## Like histograms in this case, but rotated and centered horizontally  

\newpage

```{r p22,fig.width = 25, fig.height= 30}
p22<-ggplot(leDiff_2019,aes(x=reorder(Country.Name,AvgDiff)))+
  geom_col(aes(y=AvgDiff,fill=pos))+
  coord_flip()+
  ggtitle("Life Expectancy at Birth\nCompared to the Global Average, 2019")+
  xlab("")+
  ylab("Years")+
  theme(legend.position = "none")
p22
```
## Focus on how each country compares to the rest of the world overall

\newpage

```{r saveToFile,eval=FALSE}
ggsave(plot = p2, width = 25, height = 30, dpi = 300, filename = "p2.pdf",limitsize=F)
ggsave(plot = p3, width = 25, height = 30, dpi = 300, filename = "p3.pdf",limitsize=F)
ggsave(plot = p4, width = 25, height = 30, dpi = 300, filename = "p4.pdf",limitsize=F)
ggsave(plot = p5, width = 20, height = 40, dpi = 300, filename = "p5.pdf",limitsize=F)
ggsave(plot = p6, width = 10, height = 40, dpi = 300, filename = "p6.pdf",limitsize=F)
ggsave(plot = p9, width = 20, height = 40, dpi = 300, filename = "p9.pdf",limitsize=F)
ggsave(plot = p11, width = 20, height = 40, dpi = 300, filename = "p11.pdf",limitsize=F)
ggsave(plot = p13, width = 20, height = 40, dpi = 300, filename = "p13.pdf",limitsize=F)
ggsave(plot = p14, width = 10, height = 40, dpi = 300, filename = "p14.pdf",limitsize=F)
ggsave(plot = p17, width = 20, height = 20, dpi = 300, filename = "p17.pdf",limitsize=F)
ggsave(plot = p18, width = 20, height = 20, dpi = 300, filename = "p18.pdf",limitsize=F)
ggsave(plot = p21, width = 20, height = 40, dpi = 300, filename = "p21.pdf",limitsize=F)
ggsave(plot = p22, width = 10, height = 40, dpi = 300, filename = "p22.pdf",limitsize=F)
```

\newpage

 - These are more sketches than they are finished graphics.
 - Data often has a lot — sometimes too much — to say.
 - The right way

    - Ask the data questions.
    - Start with the visualization basics.
    - Focus.
    - Iterate
    - Practice
    - let the data speak



